Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants

### 2017-01-18 cav2017/cav2017.tex
ciao
Stability test

Models \eqref{eq:ode} and \eqref{eq:plant} depend on external non-determinism in the form of input signals $u (t), u_k$, respectively. 
Feedback can be employed to affect the properties and behaviours of the process. 
%reduce the effect of parametric uncertainties and attenuate the effects of noise and disturbance. 
The most basic feedback architecture is the state feedback, 
where the control action $\vec{u}_k$ is computed by: 
%
\begin{equation}
u_k = r_{k} - K x_k, 
\end{equation}
%
where $K \in \mathbb{R}^{m \times n}$ is a feedback gain matrix, 
and $r_{k}$ is a reference signal.   
%
The closed-loop model then takes the form 
\begin{align}
x_{k+1} = ( A_d - B_d K ) + B_d r_k. \nonumber
\end{align}
The gain matrix $K$ can be set so that the closed-loop dynamics are shaped as desired, 
according to a desired goal or objective. 
A standard objective is to obtain stable closed-loop dynamics.

@@

Jury's criterion is a mathematical method used to verify whether the roots 
of a characteristic polynomial are within the unit circle by exploring solely the coefficients 
of the polynomial. The first step is to create a matrix 
$\mat{M} \in \mathbb{R}^{(n+1) \times (n+1)}$ with the first row elements 
$m_{0,j}=c_j$, where $j \in [0,n]$, $k=n-(i-1)$ and 
$c_j$ are the coefficients of the original polynomial. 
%
\begin{equation}
\left[\begin{array}{ccccc}
c_0&c_1&\hdots&c_{n-1}&c_n\\
m_{\scriptscriptstyle 0,0}-m_{\scriptscriptstyle 0,n}^*\frac{m_{0,n}^*}{m_{0,0}}&m_{\scriptscriptstyle 0,1}-m_{\scriptscriptstyle 0,n-1}^*\frac{m_{0,n}^*}{m_{0,0}}&\hdots&m_{\scriptscriptstyle 0,n-1}-m_{\scriptscriptstyle 0,1}^*\frac{m_{0,n}^*}{m_{0,0}}&0\\
\vdots&\vdots&\vdots&\vdots&\vdots\\
m_{\scriptscriptstyle i-1,0}-m_{\scriptscriptstyle i-1,k-j}^*\frac{m_{i-1,k}^*}{m_{i-1,0}}&m_{\scriptscriptstyle i-1,1}-m_{\scriptscriptstyle i-1,k-j}^*\frac{m_{i-1,k}^*}{m_{i-1,0}}&\hdots&0&0\\
m_{\scriptscriptstyle n-1,0}-m_{\scriptscriptstyle n-1,1-j}^*\frac{m_{n-1,1}^*}{m_{n-1,0}}&0&\hdots&0&0\\
\end{array}\right]
\end{equation}

Notice that every row has a polynomial one degree lower than the previous. 
The sufficient and necessary conditions for the polynomial in row $0$ 
to have all of its roots inside the unit circle are~(\cite{astrom1997computer}):
%
\begin{align}
|c_n|<c_0&&
\sum_{i=0}^n c_i > 0&&
(-1)^n\sum_{i=0}^n c_i (-1)^{n-i}> 0&&
m_{i,0}>0
\end{align}

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\section{Counterexample-Guided Abstractly-Refined Inductive Synthesis of Controllers for LTI Systems} 
\label{sec:CEGARIS} 
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

>>>>

Models \eqref{eq:ode} and \eqref{eq:plant} depend on external non-determinism in the form of input signals $u (t), u_k$, respectively. 
Feedback can be employed to affect the properties and behaviours of the process. 
%reduce the effect of parametric uncertainties and attenuate the effects of noise and disturbance. 
The most basic feedback architecture is the state feedback, 
where the control action $\vec{u}_k$ is computed by: 
%
\begin{equation}
u_k = r_{k} - K x_k, 
\end{equation}
%
where $K \in \mathbb{R}^{m \times n}$ is a feedback gain matrix, 
and $r_{k}$ is a reference signal.   
%
The closed-loop model then takes the form 
\begin{align}
\label{eq:closedloopss}
x_{k+1} = ( A_d - B_d K ) + B_d r_k.
\end{align}
The gain matrix $K$ can be set so that the closed-loop dynamics are shaped as desired, 
according to a desired goal or objective. 
A standard objective is to obtain stable closed-loop dynamics.

@@

A discrete-time system is asymptotically stable if and only if all the roots 
of the characteristic polynomial ({\it i.e.}, the eigenvalues) are inside of 
the unity circle, {\it i.e.}, their absolute value are less than 
one~\cite{astrom1997computer}. For a closed-loop system as~\eqref{eq:closedloopss}, 
the characteristic polynomial $P(z)$ should be computed as follows:
\begin{equation}
P(z)= \det( z I_{n} - A_d + B_d K ),
\end{equation}
where $I_{n}$ is the n-th order identity matrix.

Jury's criterion provides necessary and sufficient conditions for the eigenvalues 
to be within the unit circle and, consequently, the system is stable.  Given a 
polynomial $P(z)$ of the form
$$
P(z) = a_{0}z^{N} + a_{1}z^{N-1} + ... a_{N-1}z + a_{N} = 0, a_{0}\neq 0,
$$
Jury's criterion explores solely the coefficients ($a_{0},a_{1},...,a_{N-1}$) 
of $P(z)$ for checking the validity of the stability property $\phi_{stability}$.

In particular, Jury stability test is already explained in the control system 
literature ({\it e.g.}~\cite{fadali}). This study, however, limits itself to 
explain the encoding of Jury's criterion for the formal synthesis purpose. 
For the stability  test procedure, the following Jury matrix 
$M=[m_{ij}]_{(2N-2)\times N}$ is built  from $S(z)$ coefficients:
$$
M=\left[
\begin{matrix}
  V^{(0)} \\
  V^{(1)} \\
  \vdots \\
  V^{(N-2)}
 \end{matrix}
\right]\mbox{,}
$$
\noindent where $V^{(k)}=[v^{(k)}_{ij}]_{2\times N}$ such that
$$
v^{(0)}_{ij}=\begin{cases} 
a_{j-1}, & \mbox{if } i=1 \\   v^{(0)}_{(1)(N-j+1)}, & \mbox{if } i=2 
\end{cases}\mbox{, and} 
$$
$$
v^{(k)}_{ij}=\begin{cases} 
0, & \mbox{if } j>n-k\\
v^{(k-1)}_{1j}-v^{(k-1)}_{2j}\cdot\frac{v^{(k-1)}_{11}}{v^{(k-1)}_{21}}, & \mbox{if } j\leq n-k  \mbox{ and } i=1 \\
v^{(k)}_{(1)(N-j+1)}, & \mbox{if } j\leq n-k \mbox{ and } i=2 \\
\end{cases} \mbox{,}
$$
\noindent where $k\in\mathbb{Z}$, such that $0<k<N-2$. $S(z)$ is the 
characteristic polynomial of a stable system if and only if the following 
four propositions hold:
\begin{itemize}
\item $R_{1}$: $S(1)>0$;
\item $R_{2}$: $(-1)^{N}S(-1)>0$;
\item $R_{3}$: $\vert{a_{0}}\vert <a_{N}$;
\item $R_{4}$: $m_{11}>0\iff m_{31}\wedge m_{51}\wedge \dots \wedge m_{(2N-3)(1)}$.
\end{itemize}

The stability property $\phi_{stability}$ is then encoded as follows:
$$
\phi_{stability}\iff \ (R_{1} \wedge R_{2} \wedge R_{3} \wedge R_{4}).
$$

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\section{Counterexample-Guided Abstractly-Refined Inductive Synthesis of Safe Controllers for LTI Systems} 
\label{sec:CEGARIS} 
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
